Hysteresis in human computation:\\ how one task affects another

### 2014-11-19 docs/publications/2014.09_SCIENCE/new_man.tex
work on manuscript and plots

% Use only LaTeX2e, calling the article.cls class and 12-point type.

@@

\title{Computational hysteresis:\\ inter-task effects in human computation}

@@

Microtask platforms combine the efforts of large numbers 
of people to perform tasks that are difficult to automate with computers 
alone.  These platforms support near-real-time task completion, mimicking a
compute server, and fundamentally shift the role of the human from being a
consumer to being a producer of computing resources.  To realize its full
potential, we must understand how human computation differs from machine 
computation, and in particular address the biases known to be inherent in 
human cognition.  Here we show that when a worker completes a series of 
microtasks, the responses in later tasks are strongly influenced by the 
content of earlier ones.  Our findings are not solely cautionary. We find
evidence that these \textit{inter-task effects} can be
harnessed to improve the quality of responses, and help to achieve 
expert-level judgments, to increase the scope and reliability of citizen 
science and crowdsourcing initiatives.
\end{sciabstract}

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible for requesters to submit batches of small tasks to a large pool of 
workers, who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Originally used to distribute clerical work, these platforms 
increasingly serve as a means to engage experimental 
participants in a research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results \cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}

These platforms make it possible to tightly integrate human and 
machine computation.  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Ongoing research pursues the establishment of an HPU instruction set
including basic operators like \texttt{vote}, \texttt{subdivide}, \texttt{map},
\texttt{reduce}, \texttt{sentiment}, and \texttt{image\_tag}.  As was 
done with GPU programming, libraries are being developped to give the 
programmer more seamless access to HPU resources 
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  The availability of human computing platforms makes whole new classes
of problems tractible for the programmer.  For example ***.

But the ability to program with high-level human instructions comes with a 
cost: HPU output is inherently noisy.  Part of this noise could be 
explained by the random differences between workers.  Meanwhile, systematic 
cognitive biases also play a role.  
The phenomenon we attend to here represents yet
another dimension of HPU variance, and arises from the fact that a worker's 
performance changes as a function of the content of previous tasks.

There has been considerable investigation into the factors that affect the 
output from microtask work.  These include such factors as the 
level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using an alternative problem discription
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
inter-task effects on microtask platforms.

It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.  Thus, a worker's response during
one task may depend in part on previous ones.  Such \textit{inter-task} effects
would amount to a kind of \textit{hysteresis}, meaning that HPU outputs are not
only a function of the current input, but also of the history of inputs, with
important implications for crowdsourcing initiatives.

Here we describe a series of experiments using the Amazon Mechanical
Turk (MTurk) microtask platform that studied inter-task effects
for image-labeling microtasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers performed a series of image labeling tasks, 
and the first five images
they labeled were varied depending on the experimental group to which they
were assigned.  We studied the effects that changing the initial tasks 
had on the labels provided in the final tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}.
In the framing treatments, before workers begin the tasks, we told them that 
the study was funded by a lab with a particular area of interest, for example,
in the visual perception of food in one case, or the perception of tools in
another.

Surprisingly, inter-task effects were much stronger than framing.  
Our results demonstrate that initial tasks can change 
workers' focus in later tasks.  We also found that inter-task effects altered 
the level of specificity, and the diversity of vocabulary that workers used,
in ways that were not reproduced using framing.
Contrary to our initial expectations, our findings suggest that by carefully
designing workflows, the HPU programmer can leverage inter-task effects to
direct focus, tune specificity, and activate worker expressivity.  Our 
findings carry two lessons for the task designer: inter-task
effects can severely bias HPU output if HPU hysteresis is not taken into 
account, but at the same time, inter-task effects provide another technique
for improving HPU output and may help elicit expert-level annotation.

\paragraph*{Measuring computational hysteresis.}
In setting out to measure the computational hysteresis induced by priming, 
it is important to remember that we are only interested in the practical 
impact of priming on the execution of HPU algorithms.  
The distinction between measuring priming and measuring computational 
hysteresis is subtle, but it is crucial to ensure that our findings are of 
practical significance to crowdsourcing applications.

This distinction dictates our choice of tools for detecting HPU hysteresis.
A common practice for demonstrating that one population is different from
another (e.g., HPUs exposed to one prime vs another), is to use a 
$\chi^2$ test. This would suffice to demonstrate \textit{whether} priming of 
HPUs had occured, however it says nothing about the \textit{extent} of 
priming, and in particular, does not demonstrate the extent to which primed 
HPUs would execute an algorithm differently.

We opt instead to measure HPU hysteresis using
machine learning algorithms (classifiers), that have been trained to 
distinguish HPUs that were subjected to one prime, from those subjected to 
another.  If a classifier
can distinguish HPUs subjected to different primes, then it serves as 
a kind of certificate, proving 
(up to statistical confidence) that there exists 
an algorithm whose execution is affected by the priming of individual HPUs
(namely, the classifier itself).  This demonstrates that, for some algorithms,
using HPUs primed one way vs another will lead an algorithm to follow a 
different branch of execution with a certain frequency.
Statistically, our approach is grounded in the 
statistical divergence metric called L1-distance 
(alternatively \textit{total variational distance}), via the following 
inequality:
\begin{equation}
	D_\mathrm{L1} \geq 2 \eta - 1, 
	\label{l1}
\end{equation}
where $D_\mathrm{L1}$ is the L1-distance, and $\eta$ is the accuracy of
any binary classifier during validation (provided mild conditions are met). 
Since a classifer's performance estimates
a lower bound on the L1 distance, we shall denote 
$D_\mathrm{L1}^-\equiv 2\eta-1$.
The value of $D_\mathrm{L1}$ ranges from 0 to 1; when $D_\mathrm{L1} = 0$ HPUs 
shown one prime or another are 
indistinguishable, wheras if $D_\mathrm{L1} = 1$ priming is so complete that 
differently primed HPUs can be distinguished perfectly.
The
interested reader will find further discussion of the statistical properties
of our approach and proofs of the above claims in the supplementary material.

\paragraph{Experimental setup.}
We performed two experiments, soliciting respectively 2300, and 900 MTurk 
workers to label images depicting food, kitchen- and dinnerware, and cultural
scenes.  Both experiments had four treatments, and in all treatments of a 
given experiment, the last five ``test'' tasks performed by the workers were 
the same.  Two treatments for each experiment were used to explore inter-task 
effects, by making the initial five tasks for these treatments different.  
For example, in the first 
experiment, the test images depicted clutterd scenes of food and objects,
while ``initial'' images depicted either food only, or non-food objects 
(see Fig. 1).  Although we divide the tasks into initial tasks and test tasks,
no distinction is apparent from the perspective of the worker. Another two 
treatments in both experiments explored framing, 
by presenting the worker with language suggesting a particular research 
intent.  For example, in the first experiment, we stated 
``funded by the laborotary for the visual perception of \{ Objects and Tools 
$\vert$ Food and Ingredients \}''.  For emphasis, the frame text was presented
free of surrounding text, with the variable text in boldface and larger font.

The experiments and treatments are 
summarized in Table 1, which also introduces short names which we will use to 
identify specific experimental treatments and groupings.
An additional difference between \texttt{exp1} and \texttt{exp2} is that,
in both treatments for \texttt{task1}, workers were shown the test tasks in
one of five permutations, which allows us to analyze the dependance of 
inter-task effects on task proximity, independant of task content.
We provide a fuller account of the experimental design, and exhibit all the
task materials used, in the supplementary material.

@@

	\caption{ \footnotesize{ 
		We use the treatment names shown to refer to specific sections of 
		the data collected.  For example, \texttt{task2:food} refers to the
		treatment in the second experiment which was not exposed to a frame, 
		and whose initial tasks contained images of food (7th row).
	}}
	\label{table:1}
\end{table}

\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/tasks.pdf}
	\caption{Examples of slides used for the image labeling tasks. A) instructions shown to all workers; B) a framing slide; C) a slide from the initial set for treatment IMG:FOOD; D) a slide from the initial set for treatment IMG:OBJ; E) a slide from the test set shown in all treatments. The full set of slides used for all treatments is presented in the supplementary material.}
	\label{fig:task}
\end{figure}

@@

We assessed the strength of intertask and framing effects
by training a multinomial naive Bayes classifier to predict a worker's 
treatment (from two possibilities) based on the labels they provided 
in the test tasks (see Fig.~\ref{fig:theta}).
We found that in \texttt{task1}, whether a worker had labeled food or non-food 
objects during initial tasks, had a strong effect on their subsequent labels, 
biasing them by more than 30\%.  In comparison, the framing effect in \texttt{frame1} did 
not influence workers' labels enough for the classifier predict their treatment
with an accuracy significantly better than chance.  The results for 
\texttt{exp2} show a similar, but even more pronounced trend, with the bias
induced by intertask effects in \texttt{task2} probably exceeding 40\%.

It is worth noting that \texttt{task1} and \texttt{task2} differ in the kind
of comparison that they make.  In \texttt{task1},
the prime concepts (``food'' and ``objects'') are, in certain ways, very 
similar: both are relatively concrete, well-defined concepts, corresponding
to bounded physical entities that can be directly photographed.
But in \texttt{task2}, the prime concepts (``food'' and ``culture''), 
are very different:  culture is far more abstract and difficult to define
than food, and is not a bounded physical entity.  We will come back to this 
point, but note 
here that these pairs of prime concepts both lead to strong intertask effects
despite exploring very different conceptual dimensions.

As another point of comparison, we included additional framing-like 
treatments in \texttt{exp1} in which we attempted to increase the effect.
In the \texttt{echo} treatment pair, we used a more strongly worded frame:
``The purpose of this study is to understand the visual perception of
\{ Objects and Tools $\vert$ Food and Ingredients \}'', and then worker was
asked to reiterate back the purpose of the study, using a combo-box input.
We expected this to induce a strong effect because it signals that this detail
is of particular concern to the task requester, and more important than any of 
the instructions given beforehand.  
The overall effect of these primes was on par with the intertask 
effects in \texttt{task1} (see Fig.~\ref{fig:theta}), which shows just how
powerful intertask effects are.

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Difference in priming observed for various pairs of worker populations.
		The pairs of populations differed in terms of the initial tasks they
		performed (bars labeled by \textit{inter-t.}) or in terms of the framing
		to which they were exposed (bars labeled by \textit{frame}; see table 1 for
		details about the treatments).  The difference in priming was 
		measured based on the accuracy of a naive Bayes classifier in 
		resolving which treatment workers belonged to, as described in the 
		text.  
	}
	\label{fig:theta}
\end{figure}

\paragraph{Dynamics of intertask effects.} Following the observation that
intertask effects are strong, it is natural to ask how the effect of an early
task wanes as the worker proceeds through subsequent ones.  To investigate
this, in \texttt{task1} workers performed the test tasks in one of five 
permutations.  This allows us to average the intertask effect accross different
test tasks when they occupy the same position.  We found that the effect was
strongest for the first test image, and that a significant effect persisted
through the five test tasks (see Fig.~\ref{fig:theta}B).  This shows the 
extent to which the first 
test task re-primes workers, washing out the effect of initial tasks, but that
the effect is quite persistent.  In future work, it would be interesting to
investigate whether the first few tasks have a particularly strong priming
effect because they ``set the tone'' for the batch.

This set-up also allows us to see how much the each test tasks differ in their 
susceptibility to intertask effects, independant of their position among the
others.  There does appear to be substantial differences, but relative
position of the test task is more important (Fig.~\ref{fig:theta}C).

\paragraph{Effects on workers' direction of focus} 
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/delta_food.pdf}
	\caption{
		Difference in priming observed for various pairs of worker populations.
		The pairs of populations differed in terms of the initial tasks they
		performed (bars labeled by \textit{inter-t.}) or in terms of the framing
		to which they were exposed (bars labeled by \textit{frame}; see table 1 for
		details about the treatments).  The difference in priming was 
		measured based on the accuracy of a naive Bayes classifier in 
		resolving which treatment workers belonged to, as described in the 
		text.  
	}
	\label{fig:food}
\end{figure}

It is reasonable to expect that exposing a worker to tasks containing images
of food would increase her propensity to provide food-related labels in 
subsequent tasks.  To test whether this straightforward elicitation occured,
we used the wordnet corpus to operationalize the notion of food-related 
labels.  While the coverage of most western food-related words is quite good
we found that the names of many ethnic foods are missing, so we augmented
the corpus by crawling the world food section of the recipes website 
allrecipes.com.  We discuss the validation of our operationalization of 
food-words in detail in the supplementary material.

For \texttt{task1} and \texttt{frame1} we did find that exposing workers to 
food-oriented primes led them to increase their provision of food-related 
labels by about 2-3\% (see Fig.~\ref{fig:food}).  
In \texttt{echo} this effect was stronger
\texttt{echo:food} provided about 8\% more food-related words than 
\texttt{echo:obj}.  Indeed, the \%-increase in food labels observed for
\texttt{echo} was significantly larger than that for \texttt{taks1}, which is 
interesting
given that the overall priming effect observed for \texttt{task1} and 
\texttt{echo} were on par (c.f. Fig.~\ref{fig:theta}).
This makes sense, given that \texttt{echo:food} explicitly calls for a focus 
on food, whereas \texttt{task1:food} probably works by activating certain 
memories and concepts, with a more complex impact on labels.

In \texttt{exp2} we see a strikingly different effect: workers initially 
exposed to food-related tasks actually show fewer food-related labels for the
test tasks.  Recalling the content of the test tasks helps to explain this
observation.  The test tasks in \texttt{exp2} embed both the prime concepts
of food and culture, which we accomplished by using images of food having
very particular and identifiable cultural origins.  The initial tasks in
\texttt{task2:food} were, by comparison, culturally ambiguous, and the worker
would presumably be struck by the appearance of strong cultural overtones
in the test tasks.  On the other hand, due to the effect of negative priming,
workers in \texttt{task2:cult} might be somewhat desensitized to the cultural
content in the images, and instead be struck by the sudden appearance of food.

\paragraph{Effects on label vocabulary and specificity.}
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/vocab_specificity.pdf}
	\caption{
		Difference in priming observed for various pairs of worker populations.
		The pairs of populations differed in terms of the initial tasks they
		performed (bars labeled by \textit{inter-t.}) or in terms of the framing
		to which they were exposed (bars labeled by \textit{frame}; see table 1 for
		details about the treatments).  The difference in priming was 
		measured based on the accuracy of a naive Bayes classifier in 
		resolving which treatment workers belonged to, as described in the 
		text.  
	}
	\label{fig:specificity}
\end{figure}

We gain some insight into the induced intertask effects by looking at the
labels whose frequencies changed the most due to workers' early task exposure. 
For example, the ``food'' is the most strongly suppressed label among 
workers exposed to food-related initial tasks.  At first, this may seem 
unexpected, since, for \texttt{task1:food}, food labels are more common 
overall.  But ``food'' is the most generic possible food-related label, 
so this means that \texttt{task1:food} workers are trading their usage of
``food'' for more specific food references.

>>>>

% ULaTeX2e, calling the article.cls class and 12-point type.

@@

\title{Hysteresis in human computation:\\ how one task affects another}

@@

Microtask platforms combine the efforts of large numbers 
of people to perform tasks that are difficult to automate with computers 
alone.  These platforms support near-real-time task completion, mimicking a
compute server, and fundamentally shift the role of the human from consumer to 
producer of computing resources.  To realize its full
potential, we must understand how human computation differs from machine 
computation, and address the biases known to be inherent in 
human cognition.  The effects of various task-design variables on worker
performance have been investigated, but there has been little 
investigation into how consecutive tasks influence one another. Here we show 
that the responses in later tasks are strongly influenced by the content of 
earlier ones, an effect that is as strong or stronger than the effect of 
framing.  Our findings are not solely cautionary; we find
that these \textit{intertask effects} can be harnessed to improve the quality 
of responses, and thereby help to achieve expert-level judgments in citizen 
science and other crowdsourcing initiatives. 
As an ancillary contribution, we demonstrate the use of machine learning 
algorithms measure exposure-induced human bias, in terms of its ability to 
influence human computations. 
\end{sciabstract}

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible for requesters to submit batches of tasks to a large pool of 
workers, who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results 
\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Originally used to distribute clerical work, these platforms 
increasingly serve as a means to engage experimental participants in a 
research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
This trend has seeded considerable interest in characterizing the parameters
of task design, and validating the microtask platform methodology (\#).

Computer scientists view microtask platforms in the context of general
computation.  In this regard, the platforms enable a tight integration of 
human and machine computation.  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Ongoing research pursues the establishment of an HPU instruction set
including basic operators like \texttt{vote}, \texttt{subdivide}, \texttt{map},
\texttt{reduce}, \texttt{sentiment}, and \texttt{image\_tag} (\#).  Libraries 
are being developed to give the programmer more seamless access to HPU 
resources 
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  
As with other architectures, the HPU architecture makes entirely new classes 
of problems tractable for the programmer (\#).

But the ability to program with high-level human instructions comes with a 
cost: HPU output is inherently noisy.  Part of this noise can be explained
in terms of worker demography (\#).  But systematic cognitive biases also 
play a role (\#), a dimension of variability we investigate here.   
So far, investigation has explored such design factors as the level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using alternate problem descriptions
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
the influences that consecutive microtasks exert on one another.

It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.  Thus, a worker's response during
one task may depend in part on previous ones.  Such \textit{intertask} effects
would amount to a kind of \textit{hysteresis}, meaning that HPU outputs are not
only a function of the current input, but also of the history of inputs, with
important implications for the design of crowdsourcing initiatives.

Here we describe a series of experiments using the Amazon Mechanical
Turk (MTurk) microtask platform in which we studied intertask effects
for image-labeling tasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers were presented with a series of images (one at
a time), and had to provide five descriptive words (\textit{labels}) for each.
We varied first five images (hereafter the \textit{initial tasks})
and observed the effect this had on the workers' labels for the last five
(\textit{test tasks}).

The initial tasks were chosen to embed a certain \textit{prime concept}.
In our first experiment, our prime concepts were \textit{food} and (non-food) 
\textit{objects}.  In our second experiment, the prime concepts were 
\textit{food} and \textit{culture}.  The test tasks for the experiments used 
images that combined both prime concepts. Any given worker performed in only 
one experiment, and was randomly assigned initial tasks based on one of the 
prime concepts.  From the perspective of the worker, we made no distinction 
between the initial tasks and test tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}.
In the framing treatments, before workers begin the tasks, we told them that 
the purpose of study was to understand the visual perception of 
``food and ingredients'' in one case, or the 
perception of ``objects and tools'' in another.  We varied the language of 
the framing text (hereafter simply the \textit{frame}), and experimented with 
requiring the worker to echo the frame using a combo-box input.

We found that intertask effects were stronger than framing, unless the worker
was required to echo the frame, in which case both exposures were on par.  
Our results demonstrate that initial tasks can change 
workers' focus in later tasks.  We also found that intertask effects 
significantly altered the level of specificity, and the diversity of 
vocabulary that workers used, and did so in ways that differed from framing.
Contrary to our initial expectations, our findings suggest that the HPU 
programmer can leverage intertask effects.  By carefully designing workflows,
worker focus, specificity, and expressivity can be tuned.  Our 
findings carry two lessons for the task designer: intertask
effects can severely bias HPU output if uncontrolled, but if properly 
controlled, intertask effects might be optimized, which may be of significant 
utility in achieving expert-level output.

\paragraph*{Measuring computational hysteresis.}
Before describing the details of our experiments, we explain some 
choices made in measuring intertask and framing effects 
(hereafter collectively \textit{exposure effects}).  
It might be tempting to invoke the concept of a ``neutral frame of mind'' 
which would serve as a baseline for drawing comparisons.  But such a concept 
seems both difficult to define, and unlikely to arise.
Therefore, we instead always consider the effect of one exposure 
\textit{relative} to another.  Thus, our experiments consist of 
pairs of treatments between which we can make meaningful comparisons.

Our goal in measuring exposure effects is to show that they can actually 
alter the course of algorithms.
We could, by means of a $\chi^2$ test, reject the null hypothesis that
differently exposed HPUs have the same distribution of outputs.
However, this alone would not demonstrate a practically significant effect, in 
terms altering the execution path of algorithms.

For this reason, we develop a measure in which we train classifier algorithms 
to infer the exposure of an HPU from its outputs.
If a classifier infers an HPU's exposure with significantly 
better accuracy than provided by chance, then the HPU outputs must have 
been affected by the exposure.  
But in addition, the classifier serves as a kind of certificate, since it is 
an example of an algorithm whose path of execution is influenced by the 
exposure of 
individual HPUs.  In the supplementary material, we ground our approach 
rigorously in terms of the L1-distance metric, which is a measure of 
divergence between distributions.  The L1-distance ranges from 
0 to 1.  In our case, a value of 1 indicates that the HPU's exposure
influences its output so strongly that the exposure can be inferred from the
output without ambiguity.  A value of 0 indicates that the exposure has no
impact on the HPU's outputs.  The connection between the
L1-distance, $D_\mathrm{L1}$, and a classifier's accuracy, $\eta$, is given by:
\begin{equation}
	D_\mathrm{L1} \geq 2 \eta - 1, 
	\label{l1}
\end{equation}
Since a classifier's performance provides a lower bound on the L1-distance, 
we shall denote $D_\mathrm{L1}^-\equiv 2\eta-1$.
The L1-distance admits the following interpretation: if two distributions
$P$ and $Q$ have an L1-distance $d$, then in the worst case, the 
frequency with which a binary algorithm operating on samples
from $P$ outputs a 1, will differ from that when operating on samples from 
$Q$ by $d \times 100\%$.  
Thus, we can interpret $D_\mathrm{L1}^-$ as a kind of
lower bound for the worst-case bias induced by HPU hysteresis.
We provide further discussion of the statistical properties of our approach 
in the supplementary material.

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/images.jpg}
	\caption{
		Examples of images used in labeling tasks. Images from initial
		tasks: A) \textit{task1:food}, B) 
		\textit{task1:obj}, D) \textit{task2:food}, 
		E) \textit{task2:cult}.  Images from  test tasks: C) \textit{exp1}, 
		F) \textit{exp2}.  The full set of experimental materials is 
		shown in the supplementary material.
	}

	\label{fig:task}
\end{figure}
We performed two experiments, soliciting respectively 2300, and 900 MTurk 
workers to label images depicting food, kitchen- and dining-ware, and cultural
scenes.  In a given experiment, all workers performed the same test
tasks, the output of which was our main dependent variable.  Before performing
the test tasks, workers were subjected to an exposure, via one of 
two \textit{modalities}: either they completed a set of initial tasks, or 
were exposed to a frame, or both.  As mentioned, from the perspective of the worker, 
there was no distinction or discontinuity between initial and test tasks.

We organized the experiments into \textit{treatment pairs}.  Both treatments in a 
pair involved the same exposure modality, but differed in the
content of the exposure.  In Table 1, we summarize the various experimental
treatments, and introduce a naming convention that we use to refer to 
treatments and treatment pairs.
In the treatment pairs that studied intertask effects (\textit{task1} and
\textit{task2}), workers performed 10 image-labeling tasks, with the first 
five tasks comprising the initial tasks and the last five tasks comprising the
test tasks.

In the first experiment, the content of the exposures was based on the prime
concepts \textit{food} and \textit{objects} (throughout, ``objects'' should be
taken to imply ``non-edible objects'').  So, 
in \textit{task1:food}, the initial tasks contained images depicting food, 
while in \textit{task1:obj}, they contained images depicting objects, 
including kitchen and dining-ware (see Fig~\ref{fig:task}).  Meanwhile, 
for the exposures in \textit{frame1} we indicating that the study was
``funded by the laboratory for the visual perception of \{ Food and Ingredients
$\vert$ Objects and Tools~\}''.  To further strengthen the frame, we 
introduced another pair of framing treatments, \textit{echo}, with stronger
wording: ``The purpose of this study is to understand the visual perception
of \{ Food and Ingredients $\vert$ Objects and Tools~\}'', and required
the worker to echo the purpose of the study by selecting it among other 
options in a combo-box input.

Note that food and objects are both relatively well-defined classes of 
physical entities.  To explore a different conceptual axis, in 
\textit{exp2} we adopted the prime concepts food and \textit{culture}.
Compared to food and objects, culture is more abstract, difficult to define, 
and, rather than comprising a class of physical entities, it encapsulates
sets of beliefs, values, and practices.  Examples of images from the initial 
tasks in \textit{task2} are shown in Fig.~\ref{fig:task}, and the full 
collection of 
experimental materials is presented in the supplementary material.  We 
ensured that workers only participated in one treatment of one experiment.

@@

	\caption{ \footnotesize{ 
		We use the treatment names shown to refer to specific sections of 
		the data collected.  For example, \textit{task2:food} refers to the
		treatment in the second experiment which was not exposed to a frame, 
		and whose initial tasks contained images of food (7th row).
	}}
	\label{table:1}
\end{table}

@@

We assessed the strength of intertask and framing effects
by training a multinomial naive Bayes classifier to infer a worker's exposure
based on the labels provided in the test tasks (see Fig.~\ref{fig:theta}).
This inference was made over the two possibilities given the treatment pair
to which the worker had been assigned.
We found that in \textit{task1}, whether a worker performed food- or 
object-containing initial tasks, had a strong effect on their subsequent 
labels, introducing a bias that probably exceeds 30\%.  In comparison, the 
framing effect in \textit{frame1} did 
not influence workers' labels enough for the classifier to make predictions 
with significantly better accuracy than that afforded by chance.  The results 
for \textit{exp2} show a similar, but even more pronounced trend, with the bias
induced by intertask effects in \textit{task2} probably exceeding 50\%.

We see a comparable effect size in the \textit{echo} treatment pair.
We expected \textit{echo} to show a very strong effect because, by requiring
the worker to echo the purported purpose of the task, we signal that it
is of particular importance, more so than the instructions provided
beforehand.  The fact that the intertask effects were on par with the framing 
effects in \textit{echo} shows just how powerful intertask effects can be.

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Lower-bound estimates for the  worst-case bias introduced in image
		labeling tasks due to worker exposure effects.  The estimate was 
		based on the performance of a multinomial naive Bayes classifier,
		trained to infer an HPUs exposure. A) bias induced through various
		modalities and prime concepts (see Table 1). B) Bias induced in 
		\textit{task1}, as a function of test task position.  Error bars
		show the one-tailed 95\% confidence interval.
	}
	\label{fig:theta}
\end{figure}

\paragraph{Dynamics of intertask effects.} It is natural to expect 
intertask effects to attenuate as a worker proceeds through test tasks.  
In \textit{task1}, we permuted the test tasks, allowing us to 
measure intertask effects as a function of test task position, while factoring
out the content-dependant differences between test tasks
(see Fig.~\ref{fig:theta}B). 
We found intertask effects were most pronounced for the first test image,
dropping off substantially for the second test task, but maintaining a 
significant effect through until the 5th test task.  
This shows that the range of intertask effects extend accross at least 5 
tasks, and probably substantially more.
This makes intertask effects of greater concern, but as we will show, it also
makes intertask effects of greater potential utility.
In future work, it would be interesting to investigate whether a worker's
susceptibility to intertask effects varies accross a large batch.  For example,
It is conceivable that the first few tasks induce especially strong
or long-lasting effects, in effect ``setting the tone'' for the rest of the batch.

\paragraph{Detailed influences on worker outputs.} 
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/vocab_specificity.pdf}
	\caption{
		Detailed exposure effects on workers label vocabulary, for various
		exposure modalities and prime concepts (see Table 1).  A) Change in 
		the number of food-related labels used. B) Change in the total number 
		of \textit{unique} food-related labels used. C) Relative specificity
		of treatment pairs, when considering food-related labels.  For all
		plots, a positive bar height indicates an increase for the 
		food-exposed treatment.
	}
	\label{fig:specificity}
\end{figure}

Having observed significant effects from framing and task-exposure,
we looked in greater detail at the nature of the effects.  For instance,
does a food-oriented exposure induce workers to provide more food-related
labels?  And if so, does this explain the classifier's ability to distinguish
a worker's initial exposure?

To answer these questions, we used the wordnet corpus, augmented with names of
ethnic foods learned from the Internet.  Wordnet provides a (roughly)
hierarchical set of relationships between nouns, called hypernym-hyponym
relationships.  A hypernym is a generalization, while a hyponym is a 
specialization.  So, for example, ``bread'' is hypernym to ``rye bread'' 
and hyponym to ``baked goods''.

Using an augmented hypernym-hyponym structure of wordnet to identify nouns 
depicting food, we found that, for the experimental pairings in 
\textit{exp1}, the
food-exposed treatments did produce a significantly larger fraction of 
food-related labels (see Fig.~\ref{fig:specificity}A).  While the shift in 
focus with regard to food does help explain the classifier's ability to 
infer workers exposure, it still leaves most of this difference unexplained 
(*).
Furthermore, such a simple relationship cannot explain the observation for
\textit{exp2:task2} where we see a significant but reversed effect
(see Fig.~\ref{fig:specificity}A).
Given the difference between the food-culture and food-object axes, it should
not be surprising that the effect of exposure is different.  We must therefore
be prepared to accomodate multiple, countervailing factors which influencing 
the bulk number of references to the prime concept (food).  We will come back 
to this notion momentarily.

To explore these effects further, we looked more closely at which particular
labels are most suppressed (activated) by a worker's exposure.  Remarkably,
we found that, for all treatment pairs, ``food'' was always the most 
suppressed label among 
the food-exposed workers.  This is surprising, considering 
that the bulk incidence of food-related labels mostly \textit{increased} among 
food-exposed workers.  But taken together, the increase in references to food 
and the decrease in the incidence of ``food'' per se (which is the most generic
food-reference in the wordnet-based ontology), necessarily implies that
food-exposed workers have traded ``food'' for more specialized references to 
food.

In a crowdsourcing task, considerable effort must be given to shaping the 
vocabulary with which a worker engages a task.  To draw an example from the 
domain of citizen science, in galaxy zoo, workers catalog images of galaxies 
in terms of smoothness, and the presence of rings, dust lanes, digital 
disturbances, and so on.  Shaping this vocabulary is crucial to the reliability
of results.  In galaxy zoo, the interface contstrains and prompts the user
with this basic vocabulary.  But this would not appear to be sufficient, since
a set of positive and negative training examples are used, together with
peer-to-peer and peer-to-expert exchange in discussion areas, presumably to
refine and homogenize the vocabulary (\#).

In our set-up, where the worker's vocabulary is unconstrained, we can 
directly observe the effects that frames and initial tasks have on worker 
vocabulary.  By looking at the number of unique food-related labels that 
workers produced, we assessed the richness of vocabulary that workers used when
referring to food (see Fig.~\ref{fig:specificity}B.  We found that the 
food-exposed workers use a richer vocabulary of 
food-related words, even in cases like \textit{task2:food}, where they 
made fewer references to food overall.  
Interestingly, task-based exposure to food had a stronger enriching effect 
(for food-vocabulary) than did framing-based exposures. (* explain why vocab 
diversity is useful.)

Using the ontology of food induced by wordnet, we were able to directly
calculate the relative specificity of workers from different treatments.
To do this, we considered all the labels from one treatment
(for example \textit{task1:food}) and compared these to all the labels from 
another (\textit{task1:obj}).  For all possible comparisons, we observed 
which treatment's label was more specific more often, and normalized this
quantity as a percentage (we provide greater detail in the supplementary 
material).

We calculated the relative specificity for all treatment pairs, while 
restricting focus to 
food-related labels (see Fig.~\ref{fig:specificity}C).  We found that 
the food-exposed workers always provided food-references of substantially 
greater specificity.  Workers that had labeled images containing food 
were about 12\% more specific than those that labeled images of objects,
and about 20\% more specific than those that labeled images of cultural 
scenes.  With respect to the specificity of food references, framing effects 
were on par with intertask effects.

Taken together, the results presented in Fig.~\ref{fig:specificity} can 
be explained by a combination of priming and a context-dependant appraisal
of salience resembling \textit{negative} priming.
Negative priming occurs when a repeated stimulus, which is perceived to be
non-salient, begins to be ignored.
